### 2013-06-25 cws/emnlp_cws/relatedwork.tex

\section{Related Work}
If the Introduction section becomes too long, or we have some space left at the end, we can consider move some of the comparison stuff into this section.