### 2013-06-25 cws/emnlp_cws/conclusion.tex
created a separate dir for emnlp13 paper

\section{Conclusion}

In this paper we presented ... blah blah.
Stress the simplicity and effectiveness of our method.