### 2013-07-04 cws/emnlp_cws/conclusion.tex
more updates

\section{Conclusion}

In this paper we presented a straightforward approach to Chinese word segmentation using dual decomposition for system combination. We demonstrated that this method allows for joint decoding of existing CWS systems that performs better than either system alone on standard datasets for the task, and

We also demonstrated that our joint system produces more consistent segmentations than our baselines, a property that has potentially important downstream impacts for many Chinese NLP tasks. It remains to