### 2013-07-04 cws/emnlp_cws/experiment.tex

\section{Experiments}

In this work, we employ two baseline models --- a character-based CRF and a word-based perceptron --- and test the performance of jointly decoding these baseline systems with dual decomposition.

For our character-based CRF, we use the open-source Stanford CRF segmenter described in \cite{Tseng:2005:SIGHAN}.\footnote{http://nlp.stanford.edu/software/segmenter.shtml} In this system we use L2 regularization with a value of X %TODO: insert actual value / word this properly
and sigma set to 3.

For our word-based perceptron, we use a reimplementation of \cite{Zhang:2007:ACL}, run with 10 iterations of training.

For joint decoding with dual decomposition, we use an initial step size set to 0.1, and run for 100 iterations.

% question: what is the relationship between this dev data and that used in sighan2003/2005?
We use the same development set employed by \cite{Zhang:2007:ACL}, with Chinese Treebank sections 1-270, 400-931, and 1001-115 used as training data and sections 271-300 used as development data for tuning the hyperparameters of all three systems.

\subsection{Datasets}
Our test data is the standard SIGHAN 2003 \cite{Sproat:2003:SIGHAN} and 2005 \cite{Emerson:2005:SIGHAN} bake-off datasets. [Add a note here about why not using CTB - who figured out it was broken, how do we know not to use it?]

%List a table of the corpora names, size of training, testing, etc, similar to \cite{Sun:2010:COLING}.

%
%\begin{table}
%\centering
%\begin{small}
%\begin{tabular}{ l | l | c | c | c | c | c   }
%%\cline{2-10}
%%\hline
%% &   \multicolumn{1}{|c}{Models} & \multicolumn{1}{c}{$\lambda$} & \multicolumn{1}{c|}{No. of experts} & {P} & {R} & {F$_1$}   & \multicolumn{1}{c}{P} & \multicolumn{1}{c}{R} & \multicolumn{1}{c}{F$_1$}   \\
%\hline
%    \multicolumn{2}{c}{}  &  \multicolumn{1}{c}{\#W.T.} &  \multicolumn{1}{c}{\#W} &  \multicolumn{1}{c}{\#C.T.}    & \multicolumn{1}{c}{\#C}   &  \multicolumn{1}{c}{OOV}  \\ 
%\hline
%\multirow{4}{*}{2005}  & AS      & -  &  -  & -  &  - & 4.3   \\
%& MS      & 88K & 2.3M & 5K & 4.1M & 2.6 \\
%& PU      &  55k & 1.1M & 5K & 1.8M & 5.8  \\
%& CU      & 69K  & 1.5M& 5K & 2.4M & 7.4 \\
%\hline
%\multirow{3}{*}{2003} & AS  &  - & - & - & - & 2.2 \\
%& PU  &  - & - & - & - & 6.9 \\
%& CU  & - & - & - & - & 7.1  \\
%%\hline
%%\multirow{4}{*}{CTB}  &  Best03  &  \textbf{0.886}  & 0.875 &  \textbf{0.881} & 0.181 & 0.644  & \textbf{0.927} \\
%%& PcpTr       & 0.869 & 0.865 & 0.867 & 0.181 & 0.680  & 0.910 \\
%%& PcpTr & 0.865 & 0.871 & 0.868 & 0.181 & 0.660 & 0.910 \\
%%& DD  & 0.876 & \textbf{0.878} & 0.877 & 0.181 & \textbf{69.2} & 0.917  \\
%
%\end{tabular} 
%\caption{Results on SIGHAN 2005 and 2003 datasets. }\label{tbl:results}
%\end{small}
%\end{table}