### 2013-07-05 cws/emnlp_cws/conclusion.tex

\section{Conclusion}

In this paper we presented an approach to Chinese word segmentation using dual decomposition for system combination. We demonstrated that this method allows for joint decoding of existing CWS systems that performs better than either system alone, and further achieves the best performance reported to date on standard datasets for the task.

We also demonstrated that our joint system produces more consistent segmentations than our baselines. This property that has potentially important downstream impacts for many Chinese NLP tasks, though it remains to test this empirically.

Perhaps most importantly, our approach is straightforward to implement and does not require retraining of the underlying segmentation models used. This suggests its potential for broader applicability in real-world settings than existing approaches to combining character-based and word-based models for Chinese word segmentation.