### 2013-07-05 cws/emnlp_cws/discussion.tex

\section{Discussion}

%In this section we talked about detailed consistency analysis and maybe give some examples.

On the whole, dual decomposition produces state-of-the-art segmentations that are more accurate, more consistent, and more successful at inducing out-of-vocabulary words than the baseline systems that it combines.

\subsection{Dual decomposition convergence}
Plot the histogram of the number of iterations it takes to converge, and percentage of optimality.

\subsection{Consistency}

\cite{Chang:2008:ACL} have shown that increased segmentation consistency is correlated with better machine translation performance. Following their method for calculating the conditional entropy of a segmentation system, we see in Table [insert table] that our dual decomposition method achieves the most consistent segmentations (lowest conditional entropy) on 6 out of 7 datasets.

Consistency Results, Sighan 2003:

as crf
0.037814800237
as pct
0.038997075095
as dd
0.0372484903836
cityu crf
0.0923886605473
cityu pct
0.097496193366
cityu dd
0.0843064189868
ctb crf
0.170532711404
ctb pct
0.181993909505
ctb dd
0.161722439818
pku crf
0.0429803512049
pku pct
0.0552572297562
pku dd
0.0460509973844

Consistency Results, Sighan 2005:

as crf
0.0635314146511
as pct
0.0603263902622
as dd
0.055097582491
cityu crf
0.0649537861096
cityu pct
0.0730776631941
cityu dd
0.0618197636595
msr crf
0.073782773985
msr pct
0.0628241148726
msr dd
0.054883418641
pku crf
0.0888607888366
pku pct
0.0996634383663
pku dd
0.086635535204
%
%\subsection{Oracle}
%
%Following \cite{Sun:2010:COLING}, we run an oracle experiment to estimate the upper bound of improvement possible via system combination to further contextualize our results. 
%
%To do so, we combine our two baselines with the gold-standard segmentation. Each character in the test set is labeled with three \emph{B} or \emph{I} tags, \emph{B} when it begins a word and \emph{I} when it is word-medial or word-final, according to our two baselines and the gold standard. We then create oracle labels by majority vote: if the baselines agree, their label is used; if they disagree, the gold label is used.
%
%The results of this oracle experiment, shown in Table [insert table], show that [what do they show? DD approaches the upper bound possible?]
%
%% this is here to know what each column means in the output below
%%=== TOTAL INSERTIONS:	75
%%=== TOTAL DELETIONS:	84
%%=== TOTAL SUBSTITUTIONS:	783
%%=== TOTAL NCHANGE:	942
%%=== TOTAL TRUE WORD COUNT:	11989
%%=== TOTAL TEST WORD COUNT:	11980
%%=== TOTAL TRUE WORDS RECALL:	0.928
%%=== TOTAL TEST WORDS PRECISION:	0.928
%%=== F MEASURE:	0.928
%%=== OOV Rate:	0.022
%%=== OOV Recall Rate:	0.810
%%=== IV Recall Rate:	0.930
%Scoring Oracle Output, Sighan 2003
%for as
%	sighan2003/as.oracle	79	88	173	340	11989	11980	0.978	0.979	0.979	0.071	0.940	0.981
%for cityu
%	sighan2003/cityu.oracle	402	333	893	1628	35087	35156	0.965	0.963	0.964	0.073	0.823	0.976
%for ctb
%	sighan2003/ctb.oracle	929	1014	2504	4447	39921	39836	0.912	0.914	0.913	0.230	0.822	0.939
%for pku
%	sighan2003/pku.oracle	189	190	392	771	17194	17193	0.966	0.966	0.966	0.138	0.925	0.973
%Scoring Oracle Output, Sighan 2003
%for as
%	sighan2005/as.oracle	1876	748	2542	5166	122610	123739	0.973	0.964	0.969	0.099	0.890	0.982
%for cityu
%	sighan2005/cityu.oracle	577	371	1136	2084	40936	41142	0.963	0.958	0.961	0.075	0.839	0.973
%for msr // currently broken
%	sighan2005/msr.oracle	81	107	165	353	10987	10961	0.975	0.978	0.976	0.065	0.934	0.978
%for pku
%	sighan2005/pku.oracle	803	1664	2505	4972	104372	103511	0.960	0.968	0.964	0.124	0.924	0.965
%

\subsection{Error analysis}

Since dual decomposition is a method of joint decoding, it is liable to reproduce errors made by the constituent systems. In the example below, dual decomposition output follows the incorrect segmentation of the character-based CRF in oversegmenting the compound "sea, land, and air."

%some errors where DD picks the wrong model:
%gold: 有 亭台樓閣 流水
%CRF: 亭台 樓閣  <-- DD
%PCT: 亭台樓閣
\begin{tabbing}

\ \ \ \ \= \emph{English}\ \ \ \ \= Large-scale sea, land, and air \\ 
\> \> joint military exercises\\
\> \emph{Gold}\> 大规模 / 海陆空 / 联合 / 军演\\

\> \emph{CRF}\> 大规模 / 海 / 陆空 / 联合 / 军演\\

\> \emph{PCPT}\> 大规模 / 海陆空 / 联合 / 军演\\

\> \emph{DD}\> 大规模 / 海 / 陆空 / 联合 / 军演\\
\end{tabbing}

Nevertheless, in many cases the relative confidence of each model means that dual decomposition is capable of using information from both sources to generate a series of correct segmentations better than either baseline model alone. The example below shows a difficult-to-segment proper name comprised of common characters, which results in undersegmentation by the character-based CRF and oversegmentation by the word-based Perceptron, but our method achieves the correct middle ground.

\begin{tabbing}
\ \ \ \ \= \emph{English}\ \ \ \ \= Tian Yage's creations \\
\> \emph{Gold} \>   田雅各 / 的 / 创作 \\
\> \emph{CRF} \>  田雅各的 / 创作 \\
\> \emph{PCPT} \>  田雅 / 各 / 的 / 创作 \\
\> \emph{DD} \>  田雅各 / 的 / 创作 \\
\end{tabbing}

A powerful feature of the dual decomposition approach is that it can generate correct segmentation decisions in cases where a voting or polling-of-experts model could not, since joint decoding allows the sharing of information at decoding time. In the following example, both baseline models miss the contextually clear grammatical role of 上 (``above") and instead produce the otherwise common compound 上去 (``go up"); dual decomposition allows the model to generate the correct segmentation.

\begin{tabbing}
\ \ \ \ \= \emph{English}\ \ \ \ \= Enjoy a bit of snack food, ... \\
\> \emph{Gold} \>  享受 / 一点 / 点心 /  ，\\
\> \emph{CRF} \> 享受 / 一点点 / 心 / ，\\
\> \emph{PCPT} \> 享受 / 一点点 / 心 / ， \\
\> \emph{DD} \>  享受 / 一点 / 点心 / ，\\
\end{tabbing}
We found more than 400 such surprisingly accurate instances in our dual decomposition output.
%
%gold: 在 李鄭屋邨 附近
%CRF: 在 李鄭屋邨 附近
%PCT: 在 李鄭屋 邨 附近   <-- DD
%
%gold: 麥 父 笑說
%CRF:  麥 父 笑說
%PCT: 麥父 笑說    <-- DD
%
%gold: 天氣 熱 時 冷氣 大 開     
%CRF: 天氣 熱 時 冷氣 大 開     
%PCT: 天氣 熱時 冷氣 大 開     <-- DD