### 2013-07-05 cws/emnlp_cws/abstract.tex
abstract, results table changes

\begin{abstract}
There are two dominant approaches to the Chinese word segmentation problem: word-based and character-based models, each with respective strengths. Prior work has shown that gains in segmentation performance can be achieved from combining these two types of models; however, 
Past efforts have not provided a practical technique to allow mainstream adoption. We propose a method that effectively combines the strength of both segmentation schemes using an efficient dual-decomposition algorithm for joint inference. Our method is simple and easy to implement. Experiments on SIGHAN 2003 and 2005 evaluation datasets show that our method achieves the best reported results to date on 6 out of 7 datasets.
\end{abstract}

%%
%Word-based approaches are better at capturing longer context dependencies, and achieve higher segmentation consistency, whereas character-based approaches model the internal structures of words and can capture more out-of-vocabulary words. It is an intuitive idea to find methods to combine these approaches.