### 2013-07-05 cws/emnlp_cws/discussion.tex

\section{Discussion}

%In this section we talked about detailed consistency analysis and maybe give some examples.

On the whole, dual decomposition produces state-of-the-art segmentations that are more accurate, more consistent, and more successful at inducing out-of-vocabulary words than the baseline systems that it combines.

\subsection{Dual decomposition convergence}
Plot the histogram of the number of iterations it takes to converge, and percentage of optimality.

\subsection{Consistency}

\cite{Chang:2008:ACL} have shown that increased segmentation consistency is correlated with better machine translation performance. Following their method for calculating the conditional entropy of a segmentation system, we see in Table [insert table] that our dual decomposition method achieves the most consistent segmentations (lowest conditional entropy) on 6 out of 7 datasets.

%
%\subsection{Oracle}
%
%Following \cite{Sun:2010:COLING}, we run an oracle experiment to estimate the upper bound of improvement possible via system combination to further contextualize our results. 
%
%To do so, we combine our two baselines with the gold-standard segmentation. Each character in the test set is labeled with three \emph{B} or \emph{I} tags, \emph{B} when it begins a word and \emph{I} when it is word-medial or word-final, according to our two baselines and the gold standard. We then create oracle labels by majority vote: if the baselines agree, their label is used; if they disagree, the gold label is used.
%
%The results of this oracle experiment, shown in Table [insert table], show that [what do they show? DD approaches the upper bound possible?]
%
%% this is here to know what each column means in the output below
%%=== TOTAL INSERTIONS:	75
%%=== TOTAL DELETIONS:	84
%%=== TOTAL SUBSTITUTIONS:	783
%%=== TOTAL NCHANGE:	942
%%=== TOTAL TRUE WORD COUNT:	11989
%%=== TOTAL TEST WORD COUNT:	11980
%%=== TOTAL TRUE WORDS RECALL:	0.928
%%=== TOTAL TEST WORDS PRECISION:	0.928
%%=== F MEASURE:	0.928
%%=== OOV Rate:	0.022
%%=== OOV Recall Rate:	0.810
%%=== IV Recall Rate:	0.930
%Scoring Oracle Output, Sighan 2003
%for as
%	sighan2003/as.oracle	79	88	173	340	11989	11980	0.978	0.979	0.979	0.071	0.940	0.981
%for cityu
%	sighan2003/cityu.oracle	402	333	893	1628	35087	35156	0.965	0.963	0.964	0.073	0.823	0.976
%for ctb
%	sighan2003/ctb.oracle	929	1014	2504	4447	39921	39836	0.912	0.914	0.913	0.230	0.822	0.939
%for pku
%	sighan2003/pku.oracle	189	190	392	771	17194	17193	0.966	0.966	0.966	0.138	0.925	0.973
%Scoring Oracle Output, Sighan 2003
%for as
%	sighan2005/as.oracle	1876	748	2542	5166	122610	123739	0.973	0.964	0.969	0.099	0.890	0.982
%for cityu
%	sighan2005/cityu.oracle	577	371	1136	2084	40936	41142	0.963	0.958	0.961	0.075	0.839	0.973
%for msr // currently broken
%	sighan2005/msr.oracle	81	107	165	353	10987	10961	0.975	0.978	0.976	0.065	0.934	0.978
%for pku
%	sighan2005/pku.oracle	803	1664	2505	4972	104372	103511	0.960	0.968	0.964	0.124	0.924	0.965
%

\subsection{Error analysis}

Since dual decomposition is a method of joint decoding, it is liable to reproduce errors made by the constituent systems. In the example below, dual decomposition output follows the incorrect segmentation of the character-based CRF in oversegmenting the compound "sea, land, and air."

%some errors where DD picks the wrong model:
%gold: 有 亭台樓閣 流水
%CRF: 亭台 樓閣  <-- DD
%PCT: 亭台樓閣
\begin{small}
\begin{tabbing}
\ \ \ \ \= \emph{Gloss}\ \ \ \ \= Large-scale / sea, land, and air / \\ 
\> \> joint / military exercises\\
\> \emph{Gold}\> 大规模 / 海陆空 / 联合 / 军演\\

\> \emph{CRF}\> 大规模 / 海 / 陆空 / 联合 / 军演\\

\> \emph{PCPT}\> 大规模 / 海陆空 / 联合 / 军演\\

\> \emph{DD}\> 大规模 / 海 / 陆空 / 联合 / 军演\\
\end{tabbing}
\end{small}
Nevertheless, in many cases the relative confidence of each model means that dual decomposition is capable of using information from both sources to generate a series of correct segmentations better than either baseline model alone. The example below shows a difficult-to-segment proper name comprised of common characters, which results in undersegmentation by the character-based CRF and oversegmentation by the word-based Perceptron, but our method achieves the correct middle ground.

\begin{small}
\begin{tabbing}
\ \ \ \ \= \emph{Gloss}\ \ \ \ \= Tian Yage / 's / creations \\
\> \emph{Gold} \>   田雅各 / 的 / 创作 \\
\> \emph{CRF} \>  田雅各的 / 创作 \\
\> \emph{PCPT} \>  田雅 / 各 / 的 / 创作 \\
\> \emph{DD} \>  田雅各 / 的 / 创作 \\
\end{tabbing}
\end{small}

A powerful feature of the dual decomposition approach is that it can generate correct segmentation decisions in cases where a voting or polling-of-experts model could not, since joint decoding allows the sharing of information at decoding time. In the following example, both baseline models miss the contextually clear grammatical role of 上 (``above") and instead produce the otherwise common compound 上去 (``go up"); dual decomposition allows the model to generate the correct segmentation.
\begin{small}
\begin{tabbing}
\ \ \ \ \= \emph{English}\ \ \ \ \= Enjoy / a bit of / snack food / , ... \\
\> \emph{Gold} \>  享受 / 一点 / 点心 /  ，\\
\> \emph{CRF} \> 享受 / 一点点 / 心 / ，\\
\> \emph{PCPT} \> 享受 / 一点点 / 心 / ， \\
\> \emph{DD} \>  享受 / 一点 / 点心 / ，\\
\end{tabbing}
\end{small}
We found more than 400 such surprisingly accurate instances in our dual decomposition output.
%
%gold: 在 李鄭屋邨 附近
%CRF: 在 李鄭屋邨 附近
%PCT: 在 李鄭屋 邨 附近   <-- DD
%
%gold: 麥 父 笑說
%CRF:  麥 父 笑說
%PCT: 麥父 笑說    <-- DD
%
%gold: 天氣 熱 時 冷氣 大 開     
%CRF: 天氣 熱 時 冷氣 大 開     
%PCT: 天氣 熱時 冷氣 大 開     <-- DD