### 2013-07-06 cws/emnlp_cws/discussion.tex

\section{Discussion and Error Analysis}

On the whole, dual decomposition produces state-of-the-art segmentations that are more accurate, more consistent, and more successful at inducing out-of-vocabulary words than the baseline systems that it combines.
On the SIGHAN 2005 test set, over 99.1\% cases the DD algorithm converged within 100 iterations which gives optimality guarantee. 
In 77.4\% of the cases, DD converged in the first iteration. For the remaining cases, the number of iterations to convergence histogram is plotted in Figure~\ref{fig:histo}.

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{histogram.png}
\caption{No. of iterations till DD convergence histogram.}
\label{fig:histo}
\end{center}
\end{figure}

\paragraph{Error analysis}

Since dual decomposition is a method of joint decoding, it is liable to reproduce errors made by the constituent systems. In the example below, dual decomposition output follows the incorrect segmentation of the character-based CRF in oversegmenting the compound "sea, land, and air."

%some errors where DD picks the wrong model:
%gold: 有 亭台樓閣 流水
%CRF: 亭台 樓閣  <-- DD
%PCT: 亭台樓閣
\begin{small}
\begin{tabbing}
\ \ \ \ \= \emph{Gloss}\ \ \ \ \= Large-scale / sea, land, and air / \\ 
\> \> joint / military exercises\\
\> \emph{Gold}\> 大规模 / 海陆空 / 联合 / 军演\\

\> \emph{CRF}\> 大规模 / 海 / 陆空 / 联合 / 军演\\

\> \emph{PCPT}\> 大规模 / 海陆空 / 联合 / 军演\\

\> \emph{DD}\> 大规模 / 海 / 陆空 / 联合 / 军演\\
\end{tabbing}
\end{small}
Nevertheless, in many cases the relative confidence of each model means that dual decomposition is capable of using information from both sources to generate a series of correct segmentations better than either baseline model alone. The example below shows a difficult-to-segment proper name comprised of common characters, which results in undersegmentation by the character-based CRF and oversegmentation by the word-based Perceptron, but our method achieves the correct middle ground.

\begin{small}
\begin{tabbing}
\ \ \ \ \= \emph{Gloss}\ \ \ \ \= Tian Yage / 's / creations \\
\> \emph{Gold} \>   田雅各 / 的 / 创作 \\
\> \emph{CRF} \>  田雅各的 / 创作 \\
\> \emph{PCPT} \>  田雅 / 各 / 的 / 创作 \\
\> \emph{DD} \>  田雅各 / 的 / 创作 \\
\end{tabbing}
\end{small}

A powerful feature of the dual decomposition approach is that it can generate correct segmentation decisions in cases where a voting or polling-of-experts model could not, since joint decoding allows the sharing of information at decoding time. In the following example, both baseline models miss the contextually clear grammatical role of 上 (``above") and instead produce the otherwise common compound 上去 (``go up"); dual decomposition allows the model to generate the correct segmentation.
\begin{small}
\begin{tabbing}
\ \ \ \ \= \emph{English}\ \ \ \ \= Enjoy / a bit of / snack food / , ... \\
\> \emph{Gold} \>  享受 / 一点 / 点心 /  ，\\
\> \emph{CRF} \> 享受 / 一点点 / 心 / ，\\
\> \emph{PCPT} \> 享受 / 一点点 / 心 / ， \\
\> \emph{DD} \>  享受 / 一点 / 点心 / ，\\
\end{tabbing}
\end{small}
We found more than 400 such surprisingly accurate instances in our dual decomposition output.
%
%gold: 在 李鄭屋邨 附近
%CRF: 在 李鄭屋邨 附近
%PCT: 在 李鄭屋 邨 附近   <-- DD
%
%gold: 麥 父 笑說
%CRF:  麥 父 笑說
%PCT: 麥父 笑說    <-- DD
%
%gold: 天氣 熱 時 冷氣 大 開     
%CRF: 天氣 熱 時 冷氣 大 開     
%PCT: 天氣 熱時 冷氣 大 開     <-- DD