### 2013-07-06 cws/emnlp_cws/experiment.tex

\section{Experiments}

\begin{table*}[t]
\centering
\begin{small}
\begin{tabular}{ r | c | c | c | c | c | c | c | c | c | c  }
%\multicolumn{13}{c}{\large{SIGHAN 2005}} \\
\multicolumn{1}{c|}{} & \multicolumn{5}{c|}{AS} &  \multicolumn{5}{c}{PU} \\
\hline
\multicolumn{1}{c}{}   & \multicolumn{1}{|c}{R} &  \multicolumn{1}{c}{P}     &  \multicolumn{1}{c}{F$_1$}   &    \multicolumn{1}{c}{R$_{\mathrm{oov}}$}   &   \multicolumn{1}{c}{C$_{\mathrm{onst}}$}  & \multicolumn{1}{|c}{R}  &  \multicolumn{1}{c}{P}   &  \multicolumn{1}{c}{F$_1$}   &   \multicolumn{1}{c}{R$_{\mathrm{oov}}$}  &   \multicolumn{1}{c}{C$_{\mathrm{onst}}$} \\ 
Char-based CRF    &  95.2 &   93.6  & 94.4  &  58.9 & 0.064  &  94.6 &	95.3 & 94.9  &   77.8   & 0.089    \\
Word-based Perceptron &  95.8 & 95.0 & 95.4  & 69.5 & 0.060  &  94.1 & 95.5  & 94.8  &  76.7 & 0.099     \\ 
Dual-decomp      & 95.9 & 94.9   & {95.4}  & {67.7} & 0.055 & 94.8 & {95.7} & {95.3}  &  {78.7} & 0.086   \\
%\cline{2-9}
\multicolumn{1}{c|}{} &  \multicolumn{5}{c|}{CU} &  \multicolumn{5}{c}{MSR}  \\
\hline
\multicolumn{1}{c}{}   & \multicolumn{1}{|c}{R} &  \multicolumn{1}{c}{P}     &  \multicolumn{1}{c}{F$_1$}   &    \multicolumn{1}{c}{R$_{\mathrm{oov}}$}   &   \multicolumn{1}{c}{C$_{\mathrm{onst}}$}  & \multicolumn{1}{|c}{R}  &  \multicolumn{1}{c}{P}   &  \multicolumn{1}{c}{F$_1$}   &   \multicolumn{1}{c}{R$_{\mathrm{oov}}$}  &   \multicolumn{1}{c}{C$_{\mathrm{onst}}$} \\ 
Char-based CRF    & 94.7 & 94.0 & 94.3  & {76.1}  & 0.065 & 96.4 &   96.6 & 96.5  & 71.3 & 0.074 \\
Word-based Perceptron    & 94.3 & 94.0 & 94.2  & 71.7  & 0.073 &  97.0 &  97.2 & 97.1  & 74.6 & 0.063  \\
Dual-decomp  & {95.0} & 94.4  & {94.7}  & {75.3} & 0.062 &  {97.3} &  {97.4} &  {97.4}  & {76.0} & 0.055  \\
\end{tabular} 
\caption{Results on SIGHAN 2005 datasets. {R$_{\mathrm{oov}}$}  denotes OOV recall, and C$_{\mathrm{onst}}$ denotes segmentation consistency. }\label{tbl:results}
\end{small}
\end{table*}

We conduct experiments  on the SIGHAN 2003 \cite{Sproat:2003:SIGHAN} and 2005 \cite{Emerson:2005:SIGHAN} bake-off datasets to evaluate the effectiveness of the proposed dual-decomposition algorithm. We take the publicly available Stanford CRF segmenter \cite{Tseng:2005:SIGHAN}\footnote{http://nlp.stanford.edu/software/segmenter.shtml} as our character-based baseline model, and reproduce the perceptron-based segmenter from \newcite{Zhang:2007:ACL} as our word-based baseline model.
We adopted the development setting from \cite{Zhang:2007:ACL}, and used CTB sections 1-270 for training, and sections 400-931for development. The optimized hyper-parameters used are: $\ell_{2}$ regularization parameter $\lambda$ in CRF is set to $3$; perceptron is trained for 10 iterations with beam size 200; dual-decomposition is run to max iteration of 100 with initial step size of 0.1. Other than the standard precision (P), recall (R) and F$_1$ scores, we also evaluate segmentation consistency as proposed by \cite{Chang:2008:ACL}. They have shown that increased segmentation consistency is correlated with better machine translation performance. The consistency measure calculates the entropy of segmentation variations -- the lower the score the better.
%List a table of the corpora names, size of training, testing, etc, similar to \cite{Sun:2010:COLING}.

%
%\begin{table}
%\centering
%\begin{small}
%\begin{tabular}{ l | l | c | c | c | c | c   }
%%\cline{2-10}
%%\hline
%% &   \multicolumn{1}{|c}{Models} & \multicolumn{1}{c}{$\lambda$} & \multicolumn{1}{c|}{No. of experts} & {P} & {R} & {F$_1$}   & \multicolumn{1}{c}{P} & \multicolumn{1}{c}{R} & \multicolumn{1}{c}{F$_1$}   \\
%\hline
%    \multicolumn{2}{c}{}  &  \multicolumn{1}{c}{\#W.T.} &  \multicolumn{1}{c}{\#W} &  \multicolumn{1}{c}{\#C.T.}    & \multicolumn{1}{c}{\#C}   &  \multicolumn{1}{c}{OOV}  \\ 
%\hline
%\multirow{4}{*}{2005}  & AS      & -  &  -  & -  &  - & 4.3   \\
%& MS      & 88K & 2.3M & 5K & 4.1M & 2.6 \\
%& PU      &  55k & 1.1M & 5K & 1.8M & 5.8  \\
%& CU      & 69K  & 1.5M& 5K & 2.4M & 7.4 \\
%\hline
%\multirow{3}{*}{2003} & AS  &  - & - & - & - & 2.2 \\
%& PU  &  - & - & - & - & 6.9 \\
%& CU  & - & - & - & - & 7.1  \\
%%\hline
%%\multirow{4}{*}{CTB}  &  Best03  &  \textbf{0.886}  & 0.875 &  \textbf{0.881} & 0.181 & 0.644  & \textbf{0.927} \\
%%& PcpTr       & 0.869 & 0.865 & 0.867 & 0.181 & 0.680  & 0.910 \\
%%& PcpTr & 0.865 & 0.871 & 0.868 & 0.181 & 0.660 & 0.910 \\
%%& DD  & 0.876 & \textbf{0.878} & 0.877 & 0.181 & \textbf{69.2} & 0.917  \\
%
%\end{tabular} 
%\caption{Results on SIGHAN 2005 and 2003 datasets. }\label{tbl:results}
%\end{small}
%\end{table}