### 2013-08-17 cws/response.txt
adding author's response

Thank you for your thoughtful feedback and helpful suggestions. We
look forward to revising our paper in light of your comments.
Please allow us to give some detailed answers to Reviewer #2's comments.

> though it's not clear if the improvements over the baselines and previous state-of-the-art systems are significant
Statistical significance tests using the paired bootstrap resampling method with 1000 iterations (Efron and Tibshirani, 1993) shows that our DD algorithm is significantly better than the character-based CRF basedline in F1 score in 4 out of 4 cases on SIGHAN 2005 dataset at 95% confidence level, and is significantly better than the word-based perceptron baseline in 3 out of 4 cases (except for AS) at the same confidence level.
We will report full details in the final writeup upon acceptance.

> I realise space is at a premium but I think the introduction and sections 2.1 and 2.2 could easily be compressed to make room.
Great suggestion! We will make our notations more consistent and provide more details of the DD algorithm.

> They provide no analysis of why Sun 10 outperforms their model on the CU set. It would be interesting to see some discussion regarding this point. 
Sun 2010 did not make their code or system outputs available, therefore it is difficult to obtain a fine-grained comparison of our algorithm with his on the CU set. 
However, there are two high-level differences we observe: 1) Both our chracter-based and word-based baseline systems give numbers that are significantly lower (by more than 1% in F1) on this dataset than Sun 2010; although DD algorithm improves upon both baselines, it is bounded by the quality of the underlying systems. 
2) We did not perform any dataset-specific post-processing in any of our systems since the main focus of the paper is to demonstrate the merits of combining word and character-based systems using the DD algorithm. 
Various reports from the SIGHAN participants have suggested that tricks like number and date normalization post-processing can lead to significant improvements for particular datasets, which could be a factor for the performance difference.

> It converges given one test set but this doesn't constitute a guarantee. It's not a proof.
We apologize for the confusion, but the optimality guarantee that we were referring to is with respect to the constrained-optimization problem of the joint objective (Rush et al. ACL 2010 outlined a proof for the guarantee); we were not stating that the DD algorithm is guaranteed to give the optimal results for the respective task. We will clarify it in the final writeup.